---
title: "data analysis"
author: "Kate Seltzer and Mike Charles"
date: "4/16/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
```

##Read in data
```{r}
board_games_20 <- read_csv("C:\\Users\\kates\\OneDrive\\Documents\\data\\2020.csv")

board_games_21 <- read_csv("C:\\Users\\kates\\OneDrive\\Documents\\data\\2021.csv")

detailed_games <- read_csv("C:\\Users\\kates\\OneDrive\\Documents\\data\\games_detailed_info.csv")

bgg_15 <- read_csv("C:\\Users\\kates\\OneDrive\\Documents\\data\\bgg-15m-reviews.csv")

bgg_19 <- read_csv("C:\\Users\\kates\\OneDrive\\Documents\\data\\bgg-19m-reviews.csv")
```
##NARRATIVE
Research shows that in the first year of the pandemic, board game sales skyrocked -- according to the Guardian, in the first week of the official coronavirus lockdown, board game and puzzle sales in the UK rose 240%. 

To analyze what games were most popular at the height of the pandemic--from 2020 to 2021--we examined data sets from Board Game Geek, a website with over two million users who rate and discuss board games. 

##About the data

We've created two dataframes that contain board game rankings from 2020 and 2021. They each contain the game ranking, its average score, and the year it came out. Other data available describes the games in detail, including a brief summary of the game itself. The dataframes based on 2020 and 2021 have 10 columns and 19,330 and 21,831 rows respectively.

This data is imperfect: there's not an exact comparison between 2020 and 2021, and because we don't have access to data from the years before the pandemic, it's tough to conclude what exactly has changed (though we can probably say with some certainty that there was an uptick in gameplay based on the 13% increase in number of reviews between 2020 and 2021).  

```{r}
#clean data to get rid of random characters
#use this to determine how many games were in the top 100 in 2021 but not in 2020 

board_games_20 <- board_games_20 %>%
  mutate(key=str_to_upper(Name)) %>% 
  mutate(key=str_replace_all(key,":" , "")) %>%  
  mutate(key=str_replace_all(key," " , "")) %>% 
  mutate(key=str_replace_all(key,"()" , "")) %>%
  mutate(key=str_replace_all(key,",", "")) %>% 
  mutate(data_year =2020) %>% 
  group_by(key) %>% 
  arrange(Rank) %>% 
  head(100) %>% 
  arrange(Year)

board_games_21 <- board_games_21 %>%
  mutate(key=str_to_upper(Name)) %>% 
  mutate(key=str_replace_all(key,":" , "")) %>%  
  mutate(key=str_replace_all(key," " , "")) %>% 
  mutate(key=str_replace_all(key,"()" , "")) %>%
  mutate(key=str_replace_all(key,",", "")) %>% 
  mutate(data_year =2020) %>% 
  mutate(data_year =2021) %>% 
  group_by(key) %>% 
  arrange(Rank) %>% 
  head(100) %>% 
  arrange(Year)


```

Though the number of board games sold rose over the course of the pandemic, games that were popular in 2020 tended to be popular in 2021. Versions of Gloomhaven, Pandemic, Brass, Terraforming and Twilight Imperium all appeared in the top five most popular games, according to a dataset from Board Game Geek. 

```{r}
#use this to do side by side comparison of board games ranked between 2020 and 2021
compare_20_21 <-
  bind_rows(board_games_20,board_games_21) 


compare_20_21 <- compare_20_21 %>% 
  select(key, Rank, data_year)%>%
  pivot_wider(names_from = data_year, values_from = Rank)
```

What did people love about these games in 2020 and 2021? Based on comments by reviewers who left the top five games a score of 10, people seemed to favor strategic games that were cooperative rather than competitive in nature. Interesting storylines and replay value were also key components. In some cases, the game's artwork also contributed to a positive experience. People also seemed to like games that were relatively easy to pick up and could be played with kids or newcomers to the game.

Without a comparison point, it's hard to gauge if those factors influence game popularity generally, or if they were particularly important during the pandemic. Future analysis could attempt to answer whether cooperative games were more compelling during the pandemic in attempts to preserve family peace in a time where people were confined in the home for prolonged periods of time.

```{r}
#filter by top 10 ish games for comments (run several filters here)
gloomhaven_comments <-bgg_19 %>% 
  filter(name == "Gloomhaven" & rating == 10) %>%
  drop_na(comment) %>% 
  head(10)
#themes: best "dungeon crawler," interesting, legacy games, co-op, replay value

pandemic_comments <- bgg_19 %>% 
  filter(name == "Pandemic" & rating == 10) %>%
  drop_na(comment) %>% 
  head(10)
#themes: co-op, faster to learn than other co-op games like LotR, pacing, accessible for kids and non-gamers

brassbirmingham_comments <- bgg_19 %>% 
  filter(ID == 224517 & rating == 10) %>% 
  drop_na(comment) %>% 
  head(10)
#themes: strategic, elegant, art style, high intensity

terraformingmars_comments <- bgg_19 %>% 
  filter(ID == 167791 & rating ==10) %>% 
  drop_na(comment) %>% 
  head(10)
#themes: can play solo or with people, great map, cohesive rules, streamlined play, strategic

twilightimperium_comments <- bgg_19 %>% 
  filter(ID == 233078 & rating ==10) %>% 
  drop_na(comment) %>% 
  head(10)
#themes: "You wanna space game? This is that space game!!!", improved edition, space is really cool

```

We then switched out attention from the games themselves to the individuals playing said games and leaving reviews. We filtered all users and found the average rating they gave certain games to determine if certain users were prone to more positive or negative reviews. We also counted how many reviews each specific user gave. The average rating by all of the users who left 100 or more reviews was 10. This reality shows that these users are actually giving thought to the ratings that they give, likely basing it on a set of criteria that they put together. In order to learn more, we reached out to an individual with a lot of reviews to attempt to understand their thought process. However, the user declined to comment on the subject. We thought this would shed light upon how much thought goes into each review, why reviews were given and why users feel the need to leave comments. This could even be used as a seg-way into a general story on reviews, their intentions and the thought process. 

**some analysis about user reviews

With this set of code, we were attempting to determine the average review score left by individuals on Board Game Geek. In order to do this, we grouped each user together to further organize the data. From there, we calculated the average rating of these users and determined that the average score was 10 out of 10. This tells us that regardless of the amount of reviews individuals put in, they are most likely to give a 10. This may be due to individuals playing games, enjoying their experience, and logging on to Board Game Geek to share their pleasurable experience. 

With this set of code, we were attempting to find the top 10 reviewers, based on the total number of reviews, and determine the average rating they gave amongst their total reviews. In order to parse this information from the data, we used similar steps from the previous example. We grouped the data by specific users, then filtered to only show the top 10 reviews in descending order by the number of reviews. 

The individual with the most in the top 10 had 6471 while the indivudal with the lowest had 3992, with an average of 5187.8 reviews. Regarding ratings, the averages ranged from 5.359719 to 9.179932, averaging out to an average rating of 6.71409. This is interesting because it shows that the individuals in the top 10 must give genuine thought to their reviews, basing them on certain criteria over thousands of reviews. 

Based on the data we found, it appears that many users give a review of 10 as a default answer. This is interesting because, speaking from a personal perspective, I would be more likely to leave a negative review than a positive review. However, based on this data, individuals are likely most like to give a 10/10 review, regardless of how many reviews they leave. 


```{r}
#are users who leave positive reviews reviewing certain kinds of games? 
users_19 <- bgg_19%>%
  group_by(user) %>%
  summarise(avg_rating = mean(rating), reviews = n()) %>%
  arrange(desc(reviews)) %>% 
  head(10)

```

Analyzing this data, it became clear that there are several different interesting, engaging story ideas woven into these statistics. First, a story we would be interested in is a profile of a user leaving a prolific amount of reviews. It would be engaging for the reader to learn who such individuals are, why they are so passionate about games and reviews and how it has shaped their life. Without such stories, all reviews writers are just usernames on a screen, but that does not fill the full story. 

Secondly, this data could be used to create a trend story on the business of board games. The shifts from 2020 to 2021 could be proof of that trend, but also highlighting which games were successful and why. Clarifying these points will allow businesses to understand whats profitable and whats not, even allowing game inventors insight into what specific characteristics of certain games capativated certain audiences. 

This would become even more solid reporting through interviews with people who play games, giving added perspective to the reader.

The final story idea that comes to mind is analyzing the correlation between the pandemic, board game sales and the types of games play most during the pandemic. While assumptions about why board games became popular seem clear, it would beneficial to support these ideas with facts. In order to do so, one could analyze what aspects of certain games were so captivating and what that tells us possibly about human nature. All three of these stories idea could be very insightful, especially when sourced properly. 



