---
title: "Class Reference"
author: "Student name"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Consider this a personal guide to the commands and functions you will learn. In general, when you come across an R command or function that you want to remember, put it in here along with a description of what it does and when you'd use it.

## Notes 1/7

```{r}
setwd("~/GitHub/data_journalism_2022_spring")
```

```{r}
install.packages("tidyverse")
library(tidyverse)
```
##count loans, arrange most to least
ppp_maryland_loans %>%
  group_by(project_county_name) %>%
  summarise(
    count_loans = n()
  ) %>%
  arrange(desc(count_loans))
  
##notes 2/1
```{r}
maryland_ppp %>%
  select(loan_number, amount, payroll_proceed) %>%
  #mutate(name_of_column = value)
  mutate(
  percent_payroll = payroll_proceed/amount
)
```

```{r}
#ordering
maryland_ppp %>%
  select(loan_number, amount, payroll_proceed) %>%
  mutate(
  percent_payroll = (payroll_proceed/amount)*100
)  %>% arrange(desc(percent_payroll))
```



```{r}
#introduction to contents of dataset
head(maryland_ppp)
summary(maryland_ppp)
colnames(maryland_ppp)
glimpse(maryland_ppp)
```
  
```{r}
# case_when example
maryland_jobs_categories <- maryland_ppp %>%
mutate(
  jobs_retained_category = case_when(
    jobs_retained < 10 ~ 'under_10',
    jobs_retained >= 10 & jobs_retained < 25 ~ '10_to_25',
    jobs_retained >= 50 & jobs_retained < 100 ~ '50_to_100',
    jobs_retained >= 100 & jobs_retained < 250 ~ '100_to_250',
    jobs_retained >= 250 & jobs_retained < 500 ~ '250_to_500',
    jobs_retained == 500 ~ '500'
  )
)
```
##notes 2/8
```{r}
##save as itself, make id from number into character
ppp_maryland_loans <- ppp_maryland_loans %>%
  mutate(id = as.character(id))

glimpse(ppp_maryland_loans)
```

```{r}
##how to find a column where the value is NA (special value, if you use == it'll treat it as a character)
ppp_maryland_loans %>%
  group_by(forgiveness_date) %>%
  summarise(
    count=n()
  ) %>%
  filter(is.na(forgiveness_date)) %>%
  arrange(desc(count))
```

```{r}
##create new column to extract month and year from date_approved
ppp_maryland_loans %>%
  mutate(month_year_approved = floor_date(date_approved, "month")) %>%
  group_by(month_year_approved) %>%
   summarise(
    count=n()
  ) %>%
  arrange(month_year_approved)
```

```{r}
##find largest and smallest loans
ppp_maryland_loans %>%
  summarise(max_amount = max(amount),
            min_amount= min(amount))
```

```{r}
##janitor: cleans up code
library(tidyverse)
library(janitor)
```

```{r}
# standardize column names
# cleaning function
cleaned_arnold_md_loans <- arnold_md_loans %>%
  clean_names()

# display the cleaned dataset
cleaned_arnold_md_loans
```


```{r}
#rename cleaned columns
# cleaning function
cleaned_arnold_md_loans <- arnold_md_loans %>%
  clean_names() %>%
  rename(id = x1_id)

# display the cleaned dataset
cleaned_arnold_md_loans
```

```{r}
#change column data type to number
cleaned_arnold_md_loans <- arnold_md_loans %>%
  clean_names() %>%
  rename(id = x1_id) %>%
  mutate(amount = as.numeric(amount))


# display the cleaned dataset
cleaned_arnold_md_loans
```

```{r}
# check for duplicate data
cleaned_arnold_md_loans %>%
  get_dupes()
```

```{r}
# get rid of duplicate rows
cleaned_arnold_md_loans <- arnold_md_loans %>%
  clean_names() %>%
  rename(id = x1_id) %>%
  mutate(amount = as.numeric(amount)) %>%
  distinct()


# display the cleaned dataset
cleaned_arnold_md_loans
```

```{r}
#standardize zip codes

cleaned_arnold_md_loans <- arnold_md_loans %>%
  clean_names() %>%
  rename(id = x1_id) %>%
  mutate(amount = as.numeric(amount)) %>%
  distinct() %>%
  mutate(zip = str_sub(zip, start=1L, end=5L))
#standardize capitalization
  mutate(city = str_to_title(city))
# fix misspellings
  mutate(city = case_when(
    city == "Anold" ~ "Arnold",
    TRUE ~ city
  ))

# display the cleaned dataset
cleaned_arnold_md_loans
```

