---
title: "Reverse Engineering Project"
author: "Kate Seltzer and Mike Charles"
date: "March 5"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, [Over a Dozen Black and Latino Men Accused a Cop of Humiliating, Invasive Strip Searches. The NYPD Kept Promoting Him.](https://www.propublica.org/article/over-a-dozen-black-and-latino-men-accused-a-cop-of-humiliating-invasive-strip-searches-the-nypd-kept-promoting-him)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(stringr)

```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
# Load required data

# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

# Clean required data and prepare for analysis if needed. 
ccrb_complaints <- read_csv("data/allegations.csv")


```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: “A dozen of those high-ranking officers have three or more complaints with substantiated allegations; 10 are white. Five have reached the rank of deputy inspector, the lowest promoted at the commissioner’s discretion; all are white.”
* **Analysis summary**: This is a little confusing because the data seems to show that there are 24 officers with ranks above Captain who had at least three substantiated complaints. If you first filter for just Chiefs and other ranks, you get nine who had more than three substantiated complaints, and of those nine six are white. That would indicate that we're missing three of these officers (whose complaints could have been handled by the judicial system, etc). Similarly, there appears to be 11 deputy inspectors with three or more substantiated complaints, one of whom is Black and one of whom is Hispanic. I'm not sure what accounts for the discrepancy.

```{r}
# Put code to reverse engineer sentence here
ccrb_complaints %>% 
filter(rank_now =="Deputy Inspector") %>% 
filter(str_detect(board_disposition, "^Subst")) %>%
group_by(first_name, last_name, mos_ethnicity) %>% 
summarise(count=n())%>% 
arrange(desc(count))
```


```{r}
ccrb_complaints %>% 
filter(rank_now =="Chiefs and other ranks") %>% 
filter(str_detect(board_disposition, "^Subst")) %>%
group_by(first_name, last_name, mos_ethnicity) %>% 
summarise(count=n())%>% 
arrange(desc(count))
```


```{r}
# Display results of code below this codeblock

```

### Sentence 2

* **Sentence text**: “But many of the officers also said that it’s possible to be an effective cop without so many complaints. An overwhelming majority of the city’s 36,000 officers do not have a single substantiated complaint on their records; only 1 in 9 does.”
* **Analysis summary**: This shows that 3,958 officers have at least one substantiated complaint. The NYPD says they still have (and had at the time) about 36,000 officers, and 3,958/36,000 is roughly equivalent to 1/9, so this holds up.

```{r}
# Put code to reverse engineer sentence here
ccrb_complaints %>% 
filter(str_detect(board_disposition, "^Subst")) %>%
group_by(first_name, last_name) %>% 
summarise(count=n())%>% 
arrange(desc(count)) 

(3958/36000)*100
  
1/9
# Display results of code below this codeblock

```

### Sentence 3

* **Sentence text**: “Eighty-six of the roughly 420 officers in the department who currently hold a rank above captain — running precincts and other large commands and overseeing hundreds of officers — have tallied at least one misconduct allegation that was substantiated by the CCRB, meaning that investigators amassed enough evidence of offenses, ranging from bad language to pistol whippings, to say that they happened and broke patrol guidelines. The most common involved improper property entries and searches.”
* **Analysis summary**: The data here shows that there are 81 officers who are ranked above captain and have at least one substantiated complaint. The remaining 5 may not have been dealt with by the CCRB.

```{r}
# Put code to reverse engineer sentence here
ccrb_complaints %>% 
filter(rank_now =="Chiefs and other ranks"|rank_now=="Deputy Inspector"|rank_now=="Inspector") %>% 
filter(str_detect(board_disposition, "^Subst")) %>%
group_by(first_name, last_name) %>% 
summarise(count=n())%>% 
arrange(desc(count))
# Display results of code below this codeblock

```

### Sentence 4

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 5

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

-30-